---
title: "zamzamnew"
author: "Zamzam"
date: '2023-10-19'
output: html_document
---
In section 6 
1. Write a code to create a list, populate it with data (integers from 1 to 9, the first ten letters
from the English alphabet, and the words cat, dog, rabbit in factor mode). Check the modes
in this list and the class of each of its sub elements.


```{r }
myList <- list(1:9, letters[1:10], as.factor(c("cat", "dog", "rabbit")))
myList
get_mode <- function(x) {
  if (is.factor(x)) {
    return("N/A for factors")
  } else {
    uniq_x <- unique(x)
    counts <- table(x)
    max_count <- max(counts)
    modes <- uniq_x[counts == max_count]
    return(modes)
  }
}

mode_list <- sapply(myList, get_mode)
class_list <- sapply(myList, function(x) if (is.factor(x)) "factor" else class(x))

print(mode_list)
print(class_list)
```
Write a code that creates two arbitrary vectors (vec1 = (1, 2, 3), vec2 = (4, 5, 6)), and combines them into one unified vector named all.
```{r }
vec1<-c(1,2,3)
vec2<-c(4,5,6)
combined<-c(vec1,vec2)
print(combined)
```
You can assign names to every element inside a vector as displayed below. Write a code
that displays two ways of doing so.
```{r }
vect0<-c(1,2,3,4,5,6)
names(vect0)<-c("one","two","three","four","five","six")
print(vect0)
vect0<-c(1,2,3,4,5,6)
vect1<-c(one=1,two=2,three=3,four=4,five=5,six=6)
print(vect1)
```
4. Write a code to add a scalar to every element of a vector of integers from 1 to 10.
```{r }
orgvec<-1:10
scalartoadd<-8
newvec<-orgvec+scalartoadd
print(newvec)

```
5. Write a code to create a matrix of 5 × 3 of random numbers and name the columns A,B,
and C.
```{r }
randommatrix<-matrix(runif(15), nrow = 5, ncol = 3)
colnames(randommatrix) <- c("A", "B", "C")
print(randommatrix)

```
6.Write a code to create a list consisting of the integers 1, 3, and 5, and name its elements
using the terms “one”, “three”, and “five”. Provide two ways of performing this. 
```{r }
int<-c(1,3,5)
naming<-setNames(int,c("one","three","five"))
print(naming)

 
list1<-list(one=1,two=2,five=5)
print(list1)
```
In section 7 
1. Write a code to change the numbers that are greater than 10 to zero. 
```{r }
nums <- c(12,9,8,14,7,16,3,2,9)
nums > 10
nums<-ifelse(nums>10,0,nums)
print(nums)
```
1. Write a code to create a matrix x of 4 rows and 3 columns filled with a sequence of numbers
from 1 to 12.
```{r }
x <- matrix(1:12, nrow = 4, ncol = 3)
print(x)
```
2. Write a code to refer to the element of matrix x located in the fourth row and third column.

```{r }
element <- x[4, 3]
print(element)
```
3. Write a code to select the first column of matrix x. 
```{r }
first_column <- x[, 1]
print(first_column)
```
4. Write a code to select the third and the first column of matrix x. 
```{r }
selected_columns <- x[, c(1, 3)]
print(selected_columns)


```
5. Write a code to select the second row of matrix x.

```{r }


second_row <- x[2, ]
print(second_row)
```
In section 8
1. Create a factor called fdata of the following form: 
## [1] 1 2 2 3 1 2 3 3 1 2 3 3 1
## Levels: 1 2 3
```{r }


fdata <- factor(c(1, 2, 2, 3, 1, 2, 3, 3, 1, 2, 3, 3, 1), levels = c(1, 2, 3))
print(fdata)

```

2. Convert the default levels of the factor fdata to roman numerals, such that it becomes as
follows: 
## [1] I II II III I II III III I II III III I
## Levels: I II III
```{r }

levels(fdata) <- c("I", "II", "III")
print(fdata)
```
In section 10 
1. Write a code to generate a sequence of integers between 1 to 10. 
```{r }
sequence <- seq(1, 10)
print(sequence)

```
2. Write a code to to create a vector of values from 10 to 100, in increasing order, and where
each two successive elements differ by 5. 
```{r }

vector <- seq(10, 100, by = 5)
print(vector)


```
3. Write a code to create a vector of length 10 of values from 10 to 100, in increasing order,
and where each two successive elements differ by 5.
```{r }


vector2 <- seq(10, by = 5, length.out = 10)
print(vector2)

```
In section 12 
1. What is the code used to transform a Date object (e.g. January 15, 2001) to its internal form?
```{r }

 date_str <- "January 15, 2001"
date_internal <- as.Date(date_str, format = "%B %d, %Y")
print(date_internal)

```
2. When the dates are properly read into R, what is the code used to calculate the mean, range
and the time difference between the above shown dates? 

```{r }

 
release <- c(1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.7, 1.8, 1.9, 2.0)
date <- as.Date(c("2000-02-29",
                  "2000-06-15",
                  "2000-12-15",
                  "2001-06-22",
                  "2001-12-19",
                  "2002-04-29",
                  "2002-10-01",
                  "2003-04-16",
                  "2003-10-08",
                  "2004-04-12",
                  "2004-10-04"))


rdates <- data.frame(Release = release, Date = date)


print(mean(rdates$Date))

min_date <- min(rdates$Date)
max_date <- max(rdates$Date)
range <- max_date - min_date
print(range)

attach(rdates)
rdates$Date_Difference <- c(NA, diff(Date))
detach(rdates)

print(rdates)

```
In section 13 
1. What is the code used to break the sentence ’R is a free software environment for statistical
computing’ into a sequence of words based on a space? 


```{r }

sentence <- "R is a free software environment for statistical computing"
words <- strsplit(sentence, " ")[[1]]
print(words)

```
1. What is the code used to substitute the dollar sign and comma with spaces in the given
vector.
```{r }
values <- c('$11,317.35', '$11,234.51', '$11,275.89', '$11,278.93', '$11,294.94')
values_cleaned <- gsub("[$,]", " ", values)
print(values_cleaned)

```
(a) Write a code that converts the output from table tt to a data frame.
```{r }
pets = c('dog','cat','duck','chicken','duck', 'cat','dog')
tt = table(pets)
tt
```
(c) Consider the infert dataset that comes with R, and which contains information about
the education and parity of experimental subjects. Generate a cross-tabulation showing education level per parity.
```{r }
df <- as.data.frame(tt)
print(df)
data(infert)
tt <- xtabs(~ education + parity, data = infert)
print(tt)
```
(e) Use a margin= argument of c(1,2) to add margins to both the rows and columns of tt
and save the result in a new table tt12. 
```{r }
tt12 = addmargins(tt,c(1,2))
tt12
```
In section 14 
(a) Consider the Loblolly built-in dataset in R. Provide three ways to create new variable
logheight, consisting of the logarithm of the values in the height column already
existing in the dataset.
```{r }
library(datasets)
data("Loblolly")
Loblolly <- transform(Loblolly, logheight = log(height))
head(Loblolly)

data("Loblolly")
Loblolly$logheight <- log(Loblolly$height)
head(Loblolly)
 
library(dplyr)
data("Loblolly")
Loblolly <- Loblolly %>% mutate(logheight = log(height))
head(Loblolly)
```
(b) Using the iris dataset again, write a code to to create a categorical variable called
sepalgroup, based on Sepal.Length, which will be equal to 1 for lengths less than or
equal to 5, equal to 2 for lengths between 5 and 7, and equal to 3 for lengths greater
than or equal to 7.
```{r }
data(iris)
iris$sepalgroup <- ifelse(iris$Sepal.Length <= 5, 1, ifelse(iris$Sepal.Length < 7, 2, 3))
original_group <- c(1, 2, 3, 4, 5)
group <- sample(original_group, 10, replace = TRUE)  

```
(d) Continuing with the previous example, suppose we wish to recode possible values
that group can take. For example, if group is 1 or 5, we recode it to 1, if it is 2 or 4, we
recode it to 2, and if it has any other value than 1, 5, 2, or 4, we recode it to 3. What is
the code used to achieve this using only one ifelse statement? 
```{r }
newgroup <- ifelse(group %in% c(1, 5), 1, 2)
group <- c(1, 2, 3, 4, 5) 
newgroup <- ifelse(group %in% c(1, 5), 1, ifelse(group %in% c(2, 4), 2, 3))
```
(a) Redo part (4) of Sec. 14.2 using the recode() function.
```{r }
library(dplyr)
group <- c(1, 2, 3, 4, 5)  
newgroup <- recode(group, `1` = 1, `5` = 1, `2` = 2, `4` = 2, .default = 3)
```
In section 15 
(b) Write a function that prints “Hello, world!" a number of times equal to num, where num
is an arbitrary integer. Then demonstrate your code by running it on num = 3. 
```{r }
print_hello_world <- function(num) {
  for (i in 1:num) {
    cat("Hello, world!\n")
  }
}
num <- 3
print_hello_world(num)
```
(a) Specify the command used to check the arguments of a specific function, and demonstrate your code on the lm() function in R. 
```{r }
args(lm)
args(print_hello_world )
```
In section 16
(a) What is the code used to get the mean of a list of integers from 1 to 5 using lapply()
and mean() functions? 
```{r }
my_list <- list(1, 2, 3, 4, 5)
result <- lapply(my_list, mean)
print(result)
```
(a) Write a code to create a 20 by 10 matrix of normally distributed random numbers, and
then compute the mean of each column and the sum of values in each row.

```{r }
nrows <- 20
ncols <- 10
random_matrix <- matrix(rnorm(nrows * ncols), nrow = nrows, ncol = ncols)
column_means <- colMeans(random_matrix)
row_sums <- rowSums(random_matrix)
print("Matrix of Random Numbers:")
print(random_matrix)
cat("\n")

print("Mean of Each Column:")
print(column_means)
cat("\n")

print("Sum of Values in Each Row:")
print(row_sums)
```












